---
title: "wda group 11 final project"
output: html_document
date: "2022-12-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#read in all data from 01/2021 to 10/2022
```{r}
d11 <-  read.csv("202101-divvy-tripdata.csv")
d12 <-  read.csv("202102-divvy-tripdata.csv")
d13 <-  read.csv("202103-divvy-tripdata.csv")
d14 <-  read.csv("202104-divvy-tripdata.csv")
d15 <-  read.csv("202105-divvy-tripdata.csv")
d16 <-  read.csv("202106-divvy-tripdata.csv")
d17 <-  read.csv("202107-divvy-tripdata.csv")
d18 <-  read.csv("202108-divvy-tripdata.csv")
d19 <-  read.csv("202109-divvy-tripdata.csv")
d110 <-  read.csv("202110-divvy-tripdata.csv")
d111 <-  read.csv("202111-divvy-tripdata.csv")
d112 <-  read.csv("202112-divvy-tripdata.csv")
d21 <-  read.csv("202201-divvy-tripdata.csv")
d22 <-  read.csv("202202-divvy-tripdata.csv")
d23 <-  read.csv("202203-divvy-tripdata.csv")
d24 <-  read.csv("202204-divvy-tripdata.csv")
d25 <-  read.csv("202205-divvy-tripdata.csv")
d26 <-  read.csv("202206-divvy-tripdata.csv")
d27 <-  read.csv("202207-divvy-tripdata.csv")
d28 <-  read.csv("202208-divvy-tripdata.csv")
d29 <-  read.csv("202209-divvy-tripdata.csv")
d210 <-  read.csv("202210-divvy-tripdata.csv")
```

#combine data
```{r}
df1 <-  rbind(d11,d12,d13,d14,d15,d16,d17,d18,d19,d110,d111,d112,d21,d22,d23,d24,d25,d26,d27,d28,d29,d210)
```

#install.packages("tidyverse")
#install.packages("janitor")
```{r}
library("tidyverse")
library("janitor")
library("ggplot2")
```

### data structure and data cleanning, drop all the duplicated ride id

```{r}
df1 <- df1[!duplicated(df1$ride_id),]
```

# get date info
```{r}
df1$date <- as.Date(df1$started_at)
df1$hour <- strftime(df1$started_at, "%H")
df1$month <- format(df1$date, "%b")
df1$day <- format(df1$date, "%d")
df1$year <- format(df1$date, "%Y")
df1$day_of_week <- weekdays(df1$date)
```
```{r}
df1$monthn <- format(df1$date, "%m")

```

#get member_casual value
```{r}
unique(df1$member_casual)
```
#get bike type value
```{r}
unique(df1$rideable_type)
```
#get each ride's time period
```{r}
df1$ride_length <- difftime(df1$ended_at, df1$started_at, units = "mins")
```
#change ride_lenght into numeric type 
```{r}
df1$ride_length <- as.numeric(as.character(df1$ride_length))
```

#keep ride records with time period more than 1 min 
```{r}
df1 <- filter(df1, ride_length >1)
```

#data preparation for logistic regression
#column for membership
```{r}
membership = c()
for (i in 1:length(df1[,1])) {
  if (df1$member_casual[i] == "member"){
    membership[i]= 1
  }else{
    membership[i] = 0
  }
}

```

#column for classic bike
```{r}
classic_bike1 <- c()
docked_bike1 <- c()
electric_bike1 <- c()
for (i in 1:length(df1[,1])) {
  if (df1$rideable_type[i] == "classic_bike"){
    classic_bike1[i]=1
  }else{
    classic_bike1[i] = 0
  }
}
```
#column for docked bike
```{r}
for (i in 1:length(df1[,1])) {
  if (df1$rideable_type[i] == "docked_bike"){
    docked_bike1[i]=1
  }else{
    docked_bike1[i] = 0
  }
}
```

#column for electric bike
```{r}
for (i in 1:length(df1[,1])) {
  if (df1$rideable_type[i] == "electric_bike"){
    electric_bike1[i]=1
  }else{
    electric_bike1[i] = 0
  }
}
```


```{r}
df1$electric_bike <- electric_bike1
df1$docked_bike <- docked_bike1
df1$classic_bike <- classic_bike1
df1$membership <- membership
```


#logistic regression to find relationship between ebike, classic bike and ride time
#we can see we can see that both the p-value of these variables are small.all the variable are significant.
#per increasing in electric bike ride leads to 1.71 (e^(0.5371)) increase in memberships
#per increasing in classica bike ride leads to 1.86 (e^(0.6185)) increase in memberships
```{r}

summary(glm(membership~electric_bike+classic_bike+ride_length,data=df1))

```
#logistic regression to find relationship between ebike, classic bike, ride time, corelation between bike types and ride time.
#we can see we can see that both the p-value of these variables are small.all the variable are significant.
#per increasing in electric bike ride leads to 1.84 (e^(0.6118)) increase in memberships
#per increasing in classica bike ride leads to 1.91 (e^(0.6451)) increase in memberships
```{r}

summary(glm(membership~electric_bike+classic_bike+ride_length+electric_bike*ride_length+classic_bike*ride_length,data=df1))

```



